### 2021-08-10 Baseline_Model/Create_Baseline_Model.Rmd
baseline model and paper

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Questions for Sudeshna:

1. Label symmetry - baseline model is per patient, Tanish is per sequence
2. Tanish has patients in both train and test set 
3. either we need to make structured data sequence level OR make tanish's sequence date patient level 
4. rule: if ANY sequence is labeled as cognitive impairment, do we give patient that label? 
5. How do we handle folding in neither/no? 
6. If we want to employ this strategy, all patients that get a label of cognitive impairment should instantly have a patient label of cognitive concern

Okay to leave patients in test/train but make sure labels are on patient level

Ask how long it will take to retrain, ensuring accurate test/train split

next monday APOE (and maybe COVID NAT)

```{r}
library(glmnet)
library(ROCR)
library(tidyverse)
library(data.table)
library(lubridate)

# update file paths from Mayuresh
meds<-fread("data/medications.csv")
diagnoses <- fread('data/icd_diagnosis.csv')

diagnoses[,Date := ymd(ContactDTS)]
meds[,Date := mdy(Medication_Date)]

# date time filters
#diagnoses <- diagnoses[Date >= '2018-01-01' & Date <= '2018-12-31']
#meds <- meds[Medication_Date >= '2018-01-01' & Date <= '2018-12-31']

AD_Meds <- c("Galantamine", "Donepezil", "Rivastigmine", "Razadyne", "Razadyne ER", "Aricept", "Exelon","Namenda","Memantine");

meds[,AD := grepl(paste(AD_Meds, collapse="|"),Medication,ignore.case = T)]

diagnoses[,AD := (grepl("Dementia|Alzheimer's|cognitive impairment|frontotemporal dementia",DiagnosisNM,ignore.case = TRUE) | grepl("^290|^294|^331|G30|G31|780.93",ICD9CD,ignore.case = TRUE) | grepl("^290|^294|^331|G30|G31|780.93",ICD10CD,ignore.case = TRUE)) & !grepl("hydrocephalus",DiagnosisNM,ignore.case = TRUE)];

structured_variables <- data.table(EMPI = unique(c(diagnoses$EMPI,meds$EMPI)))

structured_variables[diagnoses[AD==1,.N,by=.(EMPI)],on=.(EMPI),Dx_Count := i.N]
structured_variables[meds[AD==1,.N,by=.(EMPI)],on=.(EMPI),Med_Count := i.N]
structured_variables[is.na(structured_variables)] <- 0

## read in gold standards 
gold <- fread("data/Dementia_ReferenceStandardDataset_08292019_subset_mrn.csv")
gold[,Label := ifelse(syndromic_dx %in% c(1,2,3,4,9),1,0)]
structured_variables[gold,on= .(EMPI = empi),Label := i.syndromic_dx]

test_set_EMPIs <- fread('test_empi_0.csv')
test_set <- structured_variables[EMPI %in% test_set_EMPIs$EMPI]
structured_variables <- structured_variables[!EMPI %in% test_set_EMPIs$EMPI] 
train_set_EMPIs <- fread('train_empi_0.csv')
structured_variables <- structured_variables[EMPI %in% train_set_EMPIs$EMPI]

structured_variables[,.N,by = .(Label)]
structured_variables[,EMPI := NULL]
test_set <- test_set[,.(Dx_Count,Med_Count,Label,EMPI)]
structured_variables[,Label := ifelse(Label %in% c(1,2,3,4,9),1,0)]
test_set[,Label := ifelse(Label %in% c(1,2,3,4,9),1,0)]

```

```{r}
set.seed(5281995)
f <- as.formula(Label ~ .)
x <- model.matrix(f, structured_variables)[, -1]
y <- as.matrix(structured_variables[,3])

fit = cv.glmnet(x, y, family = "binomial",type.measure = "auc")

plot(fit)
coef(fit, s = "lambda.min")

saveRDS(fit,file = "glmnet.rds")

predict(fit, newx = x, s = "lambda.min", type = "class")

results <- data.table(Prediction = predict(fit, newx = x, s = "lambda.min", type = "class")[,1],Target = y)

results[,Correct := ifelse(Prediction == Target.Label,1,0)]
results[,sum(Correct)]

# need new x

#AUC_ROC	ACC	MCC	TP	TN	FP	FN	PPV (Precision) 	NPV	TPR (Sensitivity, Recall)	TNR (Secificity, selectivity)	F1-score

perf_auc <- performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'auc')
print(perf_auc@y.values)  # .794

# write out predictions for Ayush 
test_x <- model.matrix(f, test_set)
test_y <- as.matrix(test_set[,3])
roc_data <- data.table(Prediction = predict(fit, newx = test_x, s = "lambda.min", type = "class")[,1],Target = test_y)
roc_data[,Correct := ifelse(Prediction == Target.Label,1,0)]
roc_data[,sum(Correct)]

fwrite()
```

```{r}

sens_spec <- performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"),  as.matrix(test_set[,3])), "tpr", "fpr")

saveRDS(sens_spec, "sens_spec.RDS")
saveRDS(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"),  as.matrix(test_set[,3])), "predict_prob.RDS")
saveRDS(as.matrix(test_set[,3]), "Test_labels.RDS")

perf_acc <- performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'acc')
print(perf_acc@y.values) # .81

# TP
res <- data.table(preds = predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="class"),truth =  as.matrix(test_set[,3]))
res[,.(TP = preds.1 == '1' & truth.Label == 1)][,.N,by=.(TP)] # 18

# TN
res[,.(TN = preds.1 == '0' & truth.Label == 0)][,.N,by=.(TN)] # 43

# FP
res[,.(FP = preds.1 == '1' & truth.Label == 0)][,.N,by=.(FP)] # 0

# FN
res[,.(FN = preds.1 == '0' & truth.Label == 1)][,.N,by=.(FN)] # 16

metrics <- data.table(
cutoffs = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'ppv')@x.values[[1]], perf_ppv = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'ppv')@y.values[[1]], perf_npv = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'npv')@y.values[[1]], perf_tpr = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'tpr')@y.values[[1]], perf_tnr = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'tnr')@y.values[[1]], perf_f1 = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'f')@y.values[[1]], acc = performance(prediction(predict(fit, s=fit$lambda.min, as.matrix(test_set[,1:2]), type="response"), as.matrix(test_set[,3])),'acc')@y.values[[1]])

# do 1 minus this so mirror image
jpeg("ROC.jpeg",height=500,width=500)
plot(sens_spec,cex.axis=5)
dev.off()

fpr_tpr <-performance(prediction(predict(fit, s=fit$lambda.min, x, type="response"), y), "tpr","fpr")

cutoffs <- data.table(cut=fpr_tpr@alpha.values[[1]], fpr=fpr_tpr@x.values[[1]], 
                      tpr=fpr_tpr@y.values[[1]])

cutoffs <- cutoffs[order(-tpr)]   # 0.6402498 seems to be optimal threshold to get sensitivity equal to current test

fwrite(cutoffs,file="sens_spec_cutoffs.csv")

new_results[,.N,by=.(Target.Label)] # 452 true negatives, 452/778 = .58097

best.threshold <- .35
ux <- predict(fit,as.matrix(test_set[,1:2]),s=fit$lambda.min, type="response")
probs.new <- ifelse(ux>best.threshold,1,0)
new_results <- data.table(Predicted = probs.new[,1],Target = as.matrix(test_set[,3]))
new_results[,Correct := ifelse(Predicted == Target.Label,1,0)]
new_results[,sum(Correct)]

new_results[,EMPI := test_set$EMPI]
new_results[gold,on=.(EMPI = empi),Label := i.syndromic_dx]

fwrite(new_results,file= 'results_structured_only.csv')

# best.threshold <- 0
# ux <- predict(fit,x,s=fit$lambda.min, type="response")
# probs.new <- ifelse(ux>best.threshold,1,0)
# new_results <- data.table(Predicted = probs.new[,1],Target = y)

# new_results[Predicted == 1 & Target.Label == 0,.N]/nrow(new_results[Target.Label==0])
# 
# # values between  0.3983138 and .4001
# 
# best.threshold <- .3983138
# ux <- predict(fit,x,s=fit$lambda.min, type="response")
# probs.new <- ifelse(ux>best.threshold,1,0)
# new_results <- data.table(Predicted = probs.new[,1],Target = y)
# 
# new_results[Predicted == 1 & Target.Label == 0,.N]/nrow(new_results[Target.Label==0])

threshold_function <- function(threshold){
  best.threshold <- threshold;
  ux <- predict(fit,x,s=fit$lambda.min, type="response");
  probs.new <- ifelse(ux>best.threshold,1,0);
  new_results <- data.table(Predicted = probs.new[,1],Target = y);
  new_results[Predicted == 1 & Target.Label == 0,.N]/nrow(new_results[Target.Label==0]);
}

test_fpr <- map_dfr(c(seq(.3983138,.4001,by=.00001),.42),~data.table(Cut = .x,FPR = threshold_function(.x)))

threshold_random <- function(threshold){
  best.threshold <- threshold;
  ux <- predict(fit,x,s=fit$lambda.min, type="response");
  ux <- ux + rnorm(length(ux),.00001,.0001);
  probs.new <- ifelse(ux>best.threshold,1,0);
  new_results <- data.table(Predicted = probs.new[,1],Target = y);
  new_results[Predicted == 1 & Target.Label == 0,.N]/nrow(new_results[Target.Label==0]);
}

# number of rows that are equal to min response

responses <- data.table(Response = predict(fit,x,s=fit$lambda.min, type="response"))
responses <- responses + rnorm(nrow(responses),.0000001,.000001);

perf <- performance(prediction(responses, y),'auc')
print(perf@y.values)
sens_spec <- performance(prediction(responses, y), "tpr", "fpr")

plot(sens_spec,cex.axis=5)

```